---
title: "DSC424Problem2HomeWork4RSyntaxfile"
author: "Ronaldlee Ejalu"
date: "3/7/2021"
output:
  word_document: default
  html_document: default
---

#load the necessary packages
```{r}
library(dplyr)
library(corrplot) # correlation visualization
library(MASS) #for CV
library(caTools) # for splitting the data set into training and testing
```


#set the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC424/HomeWork4")
```

#Load the data set
```{r}
heartds <- read.csv(file="../HomeWork4/datasets/heart.csv", header=TRUE, sep=",")
```

#the top 6 rows of this data set
```{r}
head(heartds)
```

#structure of the ds
```{r}
str(heartds)
```

#check if there are any missing values
```{r}
sum(is.na(heartds))
```

#Convert all the variables to numeric
```{r}
heartdscleansed <- heartds %>%
  transmute(age = as.numeric(age)
        , sex = as.numeric(sex)
        , cp = as.numeric(cp)
        , trestbps = as.numeric(trestbps)
        , chol = as.numeric(chol)
        , fbs = as.numeric(fbs)
        , restecg = as.numeric(restecg)
        , thalach = as.numeric(thalach)
        , exang = as.numeric(exang)
        , oldpeak = oldpeak
        ,  slope = as.numeric(slope)
        , ca = as.numeric(ca)
        , thal = as.numeric(thal)
        , heartdisease =  as.factor(heartdisease)
        )
```


#structure of the heartdscleansed data set
```{r}
str(heartdscleansed)
```

#determine how heartdisease is split
```{r}
tabulate(heartds$heartdisease)
```



#Run the correlation matrix and check for VIF
```{r}
#m <- cor(heartdscleansedss)
#round(m, 2)
#m <- cor(heartdscleansedss, use='pairwise.complete.obs')
#round(m,2)
# M <- cor(heartdscleansedss)
# m <- corrplot(M, method="ellipse")
# m
# m <- corrplot(M, method="number")
# m
# model <- lm(heartdisease ~., data = heartds)
# VIF(model)
```

#########################################################################################################################################
#Running LDA with cross-validation
########################################################################################################################################

#The depedent variable must be categorical
```{r}
heartLDA <- lda(heartdisease ~., data = heartdscleansed)
#heartLDA
```   

#plotting the model
#we have to create a new without cross validation
```{r}
#heartdiseaseLDA <- lda(heartdisease ~., data=heartdscleansed)
#heartdiseaseLDA
#plot(heartdiseaseLDA, xlab = "LDA1", ylab = "LDA2")
```

# Try to predict the class from the original data
# Note ... this is JUST a test to see how this works
# In practice you will want to use cross-validation!
```{r}
p <- predict(heartLDA, newdata=heartdscleansed[,1:13])$class
p
```


#Compare the results of the prediction
```{r}
t <- table(p, heartdscleansed$heartdisease)
t
```

#determine the accuracy
```{r}
acc <- mean(p==heartdscleansed$heartdisease)
acc
```



################################################################################################################################################
#splitting the data set into training and testing
################################################################################################################################################

#you're going to do a set seed, because every set seed will give you a different randomization 
#So if you want to have the same exact.
#If you want to know exactly where you got that randomization from, you want to set the seed  to the same number
```{r}
set.seed(123)
```
#put in sample split the original data set in this case we're going to do a 7030 split 
#so 70% is going into training 
#and 30% is going into testing.
```{r}
sample <- sample.split(heartdscleansed, SplitRatio = 0.70)#Splits the data in the ratio mentioned in the SplitRatio
train <- subset(heartdscleansed, sample==TRUE)
test <- subset(heartdscleansed, sample==FALSE)
```

#The depedent variable must be categorial (Assuming No Cross-Validation)
```{r}
heartLDA <- lda(heartdisease ~., data = train)
heartLDA
#plot(heartLDA)
#ls(heartLDA)
```
#we can predict on the training
```{r}
p <- predict(heartLDA, newdata = heartdscleansed[,1:13])$class
p
```

#we can use the table to get the same confusion matrix 
```{r}
t1 <- table(p, heartdscleansed$heartdisease)
t1
```

#calculate the accuracy by using the mean
```{r}
acc1 <- mean(p==heartdscleansed$heartdisease)
acc1
```
#We can also do as we said before the Cross validation come up with the table, 
#and if we use coefficients remember if we do Ls.
# Setting "CV = T" will have the lda function perform
# "Leave-one-out" cross-validation
```{r}
heartLDA2 <- lda(heartdisease ~., data=heartdscleansed, CV=T)
heartLDA2
```

#use the table() to get a confusion matrix by using the class
```{r}
t2 <- table(heartLDA2$class, heartdscleansed$heartdisease)
t2
```

#calculate the accuracy by hand
```{r}
accuracy <- (373 + 476)/(373 + 476 + 126 + 50)
accuracy
```

#Coming up with the different performance metrics
#build a model with a train()
```{r}
library(caret)
heartModelFit <- train(heartdisease ~., method = "lda", preProcess=c("scale","center"), data=train)
```
#And then predict this data on the training data and come up with a confusion matrix.
```{r}
p <- predict(heartModelFit, train)
cm <- confusionMatrix(train$heartdisease, p, dnn=c("Actual Group", "Predicted Group"))
cm
```


