---
title: "HomeWorkAssignment3Question3"
author: "Ronaldlee Ejalu"
date: "2/21/2021"
output:
  word_document: default
  html_document: default
---

#Libraries
```{r}
library(DescTools)
library(Hmisc) #Describe Function
library(psych) #Multiple Functions for Statistics and Multivariate Analysis
library(GGally) #ggpairs Function
library(ggplot2) #ggplot2 Functions
library(vioplot) #Violin Plot Function
library(corrplot) #Plot Correlations
library(REdaS) #Bartlett's Test of Sphericity
library(psych) #PCA/FA functions
library(factoextra) #PCA Visualizations
library("FactoMineR") #PCA functions
library(ade4) #PCA Visualizations
library(yacca) # CCA
library(xlsx) # Excel
```


###################################################################
# This is a nice function for computing the Wilks lambdas for 
# CCA data from the CCA library's method
# It computes the wilkes lambas the degrees of freedom and te 
# p-values
###################################################################
```{r}
ccaWilks = function(set1, set2, cca)
{
  ev = ((1 - cca$cor^2))
  ev
  
  n = dim(set1)[1]
  p = length(set1)
  q = length(set2)
  k = min(p, q)
  m = n - 3/2 - (p + q)/2
  m
  
  w = rev(cumprod(rev(ev)))
  
  # initialize
  d1 = d2 = f = vector("numeric", k)
  
  for (i in 1:k) 
  {
    s = sqrt((p^2 * q^2 - 4)/(p^2 + q^2 - 5))
    si = 1/s
    d1[i] = p * q
    d2[i] = m * s - p * q/2 + 1
    r = (1 - w[i]^si)/w[i]^si
    f[i] = r * d2[i]/d1[i]
    p = p - 1
    q = q - 1
  }
  
  pv = pf(f, d1, d2, lower.tail = FALSE)
  dmat = cbind(WilksL = w, F = f, df1 = d1, df2 = d2, p = pv)
}
```

#Set Working Directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC424/HomeWork3")
```

#Load the data set
```{r}
responses <- read.csv(file="../HomeWork3/datasets/responses.csv", header=TRUE, sep=",")
```

#Check Sample Size and Number of Variables
```{r}
dim(responses)
```

#Show for first 6 rows of data
```{r}
head(responses)
names(responses)
```

#Check for Missing Values (i.e. NAs)
#For All Variables
```{r}
sum(is.na(responses))
```

#Listwise Deletion
```{r}
responses2 <- na.omit(responses)
```
#Check new data has no missing data
```{r}
sum(is.na(responses2))
```

#Show Structure of Dataset
```{r}
#str(responses2, list.len=ncol(responses2))
str(responses2)
```

#Anything you do to the right of the comma, you are subsettign columns
#Anything you do to the left of the comma, you are subsetting rows
```{r}
responses3 <- responses2[,c(1:73,76,77:107,110:132,134:140,141:144)]
music <- responses2[,1:19]
movie <- responses2[,20:31]
hobbies_interests <- responses2[,32:63]
phobias <- responses2[,64:73]
health <- responses2[,76]
personality_views_opinions <- responses2[,c(77:107,110:132)]
spending <- responses2[,134:140]
demographics <- responses2[,141:144]
```

# Run the CCA model
```{r}
options(scipen = 0)
c2 = cca(hobbies_interests,spending)
summary(c2)
```

#FunctionNames
```{r}
ls(c2)
```

#get the degrees of fredom
#get the chisquare
```{r}
c2$chisq
c2$df
```

#calculate the chisquare test
```{r}
round(pchisq(c2$chisq
             , c2$df
             , lower.tail=F
             )
      , 3
      )
```

# Extra credits Correspndence Analysis

# set to the working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC424/HomeWork3")
```


# Load the data file
```{r}
sport  <- read.xlsx(file="../HomeWork3/datasets/sport.xls"
                    , sheetIndex=1)
```

#first 6 rows
```{r}
head(sport)
```

#struture
```{r}
str(sport)
```

#Edit the First column as Rownames of data
```{r}
sport.colnames.rownames = sport[,1]
sport.countries = sport[, -1]
#sport.countries
row.names_scales = sport.colnames.rownames
```

#qa
```{r}
row.names(sport)
```

#convert the data into contigency table
```{r}
dtcountries = as.table(as.matrix(sport.countries))
```

#load the library ca
```{r}
library(ca)
fit = ca(dtcountries)
plot(fit)
#fits = ca(row.names_scales)
#plot(fits)
```

#
