---
title: "DSC424HomeWork1"
author: "Ronaldlee Ejalu"
StudentID: 2020637
date: "1/15/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Load all the necessary libraries 
```{r}
library(readr)
library(tidyverse)
library(gtsummary)
library(tableone)
library(broom)
library(dplyr) #dplryr calculations
library(corrplot) # Plot Correlations
library(DescTools) # VIF Function
```


# Read data into R studio
```{r}
insurance_dataset <- read_csv("C:\\Users\\rejalu1\\OneDrive - Henry Ford Health System\\DSC424\\Data Sets\\insurance_dataset.csv")
```

#view the 10 data observations
```{r}
head(insurance_dataset)
```

#view the last 10 data observations
```{r}
tail(insurance_dataset)
```

#data structure
```{r}
str(insurance_dataset)
```

# summarized statistical data from the data set
```{r}
summary(insurance_dataset)
```

#check for any missing value
#There are no missing values
```{r}
sum(is.na(insurance_dataset))
```

#get specific column index in R
```{r}
as.data.frame(colnames(insurance_dataset))
```



#distinct values of each factor column
# gender_num, smoker_num, region_num
```{r}
insurance_dataset.sex <- count(distinct(insurance_dataset), sex)
insurance_dataset.sex

insurance_dataset.smoker <- count(distinct(insurance_dataset), smoker)
insurance_dataset.smoker

insurance_dataset.region <- count(distinct(insurance_dataset), region)
insurance_dataset.region
```


# data cleaning
```{r}
insurance.clean <- insurance_dataset %>%
  transmute(age = age
            , sex = as.factor(sex)
            , gender_num = gender_num
            , bmi = bmi
            , children = children
            , smoker = as.factor(smoker)
            , smoker_num = smoker_num
            , region = as.factor(region)
            , region_num = region_num
            , expenses = expenses) %>%
  mutate(
    sex = relevel(sex, ref = 'male')
    , smoker = relevel(smoker, ref = 'no')
    , region = relevel(region, ref = 'southeast')
  )

# Create a matrix for sex
sexdummies.matrix <- model.matrix(~insurance.clean$sex)

# Convert the model matrix into a data frame
sexdummies.frame <- data.frame(sexdummies.matrix)

# bind the data frame to data set
insurance.clean <- cbind(insurance.clean, sexdummies.frame)

# create a matrix for smoker
smokerdummies.matrix <- model.matrix(~insurance.clean$smoker)

#Convert the model matrix into a data frame
smokerdummies.frame <- data.frame(smokerdummies.matrix)

#bind the data frame to data set
insurance.clean <- cbind(insurance.clean, smokerdummies.frame)

# create a matrix for region
regiondummies.matrix <- model.matrix(~insurance.clean$region)

# Convert the model matrix into a data frame
regiondummies.frame <- data.frame(regiondummies.matrix)

# bind the data frame to a data set
insurance.clean <- cbind(insurance.clean, regiondummies.frame)
```

# rename and select all the variables interest
```{r}
insurancecleansed <- insurance.clean %>%
  select(age = age
            , gender_num = gender_num
            , bmi = bmi
            , children = children
            , smoker_num = smoker_num
            , region_num = region_num
            , expenses = expenses
            , sexfemale = insurance.clean.sexfemale
            , smokeryes = insurance.clean.smokeryes
            , northeast = insurance.clean.regionnortheast
            , northwest = insurance.clean.regionnorthwest
            , southwest = insurance.clean.regionsouthwest)
```

#extract out all numerical variables
```{r}
insurance.numvariables <- insurance_dataset[,c(1,4:5,10)]
```

# check for multicollinearity amongst the numerical variables
```{r}
M <- cor(insurance.numvariables, method = "spearman")
M
corrplot(M, method = "number")
```



# Check for multicollinearity amongst all the variables
```{r}
m2 <- cor(insurancecleansed, method = "spearman")
corrplot(m2, method = "number")

```

#Altering my data set after sensing multicollinearity in the original data
# I select the variables of interest 
```{r}
insurancewithselectedvars <- insurancecleansed %>%
  select(age = age
            , bmi = bmi
            , children = children
            , expenses = expenses
            , sexfemale = sexfemale
            , smokeryes = smokeryes
            , northeast = northeast
            , northwest = northwest
            , southwest = southwest)
```

# Again check for multicollinearity
```{r}
#summary(insurancewithselectedvars)
m3 <- cor(insurancewithselectedvars, method = "spearman")
#m3
corrplot(m3, method = "number")

model2 <- lm(expenses ~ ., data = insurancewithselectedvars)
summary(model2)
VIF(model2)
```

# Explanatory analysis

#Histogram
```{r}
hist(insurancewithselectedvars$expenses, col="yellow", freq=TRUE)
x <- seq(0, 60000, length.out = 50)
y <- with(insurancewithselectedvars, dnorm(x, mean(expenses), sd(expenses)))
lines(x, y, col="red")
```


# Five - Number Summary for the Boxplot
```{r}
summary(insurancewithselectedvars$expenses)

# Boxplots
insuranxebloxplot <-ggplot(insurancewithselectedvars, aes(y=expenses)) +
  geom_boxplot(col="blue") +
  labs(
    title="expenses",
    y="expenses")
insuranxebloxplot

ggsave("insuranxebloxplot.png")
```

# Return a vector with a mean value across each row of the insurance.numvariables data set
```{r}
insurance.numvariables.means <- rowMeans(insurance.numvariables, na.rm=TRUE)
hist(insurance.numvariables.means)
```

#remove entries with the means greater than 5000
```{r}
insurance.keep <- insurance.numvariables.means < 5000
```

# remove outliers from the original data frame
```{r}
insuracedataset <- insurance_dataset[insurance.keep,]
```

# remove outliers from the numerical insurance data set
```{r}
insurance.numvariables.withoutliers <- insurance.numvariables[insurance.keep,]
```

#remove outliers from the insurance with selected vars data set
```{r}
insurancewithselectedvars.withoutliers <- insurancewithselectedvars[insurance.keep, ]
```



#plot the means with outliers removed
```{r}
insurance.mean2 <- rowMeans(insurance.numvariables.withoutliers, na.rm = TRUE)
hist(insurance.mean2)
```

#five number summary
```{r}
summary(insurancewithselectedvars.withoutliers$expenses,)
```


#box plot
```{r}
boxplot(insurancewithselectedvars.withoutliers$expenses, col = "blue", main = "Expenses", ylab = "expenses")
```


```{r}
hist(insurancewithselectedvars.withoutliers$expenses, col="yellow", freq=TRUE)
x <- seq(0, 60000, length.out = 50)
y <- with(insurancewithselectedvars.withoutliers, dnorm(x, mean(expenses), sd(expenses)))
lines(x, y, col="red")
```

# model building after removing the outliers.
```{r}
model3 <- lm(expenses ~ ., data = insurancewithselectedvars.withoutliers)
summary(model3)
```


# Creating the automatic models
```{r}
null = lm(expenses ~ 1, data = insurancewithselectedvars.withoutliers)
null

full = lm(expenses ~ ., data = insurancewithselectedvars.withoutliers)

#Forward regression
train_forward = step(null, scope = list(lower=null, upper=full),direction="forward")
summary(train_forward)
```

# using backward
```{r}
train_backward = step(full, dierction="backward")
summary(train_backward)
```

# using stepwise Regression
```{r}
train_step = step(null, scope = list(upper=full), direction = "both")
summary(train_step)
```




# Data Visualization
plot a box plot
```{r}
counts <- table(insuracedataset$smoker)
counts
barplot(counts, main="Number Of Observation per smoking status",ylab="Number Of Observations", xlab="Smoking Status", col=c("blue","green"))
```

# calculate the mean expense by smoking status
# plot a bar chart
```{r}
library(scales)
plotdata <- insuracedataset %>%
  group_by(smoker) %>%
  summarize(mean_expenses = mean(expenses))

#plotdata

# plot the means
ggplot(plotdata,
       aes(x = smoker, 
           y = mean_expenses)) +
  geom_bar(stat = "identity",
           fill = "cornflowerblue") +
  geom_text(aes(label = dollar(mean_expenses)), 
            vjust = -0.25) +
  scale_y_continuous(breaks = seq(0, 30000, 2000), 
                     label = dollar
                     ) +
  labs(title = "Mean Insurance expenses by smoking status", 
       x = "smoking status",
       y = "mean_expenses")

```























