---
title: "DSC424_Project"
author: "Kate Burns, Abigail Keller, Ronaldlee Ejalu"
date: "3/5/2021"
output: pdf_document
---

#Libraries
```{r}
library(PCAmixdata)
library(foreign)  # Allows us to read spss files!
library(corrplot)
library(QuantPsyc)
library(leaps)
library(RColorBrewer)
library(DescTools)
library(Hmisc) #Describe Function
library(psych) #Multiple Functions for Statistics and Multivariate Analysis
library(GGally) #ggpairs Function
library(ggplot2) #ggplot2 Functions
library(vioplot) #Violin Plot Function
library(corrplot) #Plot Correlations
library(REdaS) #Bartlett's Test of Sphericity
library(psych) #PCA/FA functions
library(factoextra) #PCA Visualizations
library("FactoMineR") #PCA functions
library(ade4) #PCA Visualizations
##############################################################################################
```


```{r}
# Set the working directory
setwd("~/Documents/DePaul MS DS/DePaul Winter 2021/DSC 424/Project_424")
nhanes <- read.csv(file="nhanesCompletedv4.csv", header=TRUE, sep=",")
#setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC424/HomeWork4")
#nhanes <- read.csv(file="../HomeWork4/datasets/nhanesCompletedv4.csv", header=TRUE, sep=",")
sum(is.na(nhanes))
```

############################################################################################
Principal Component Analysis
###########################################################################################

#Check for Multicollinearity with Correlations
```{r}
training = nhanes[, -c(1)]
M<-cor(training, method="spearman")
M
corrplot(cor(training,method="spearman"), method = "number", type = "lower")
#GGplot Correlation
ggcorr(training, method = c("pairwise","spearman"), label=TRUE)
ggcorr(training, method = c("pairwise","spearman"))
```

#run original PCA method with all numeric variables
```{r}
#removed most categorical and highly correlated variables (left some binary) 
training = nhanes[, c(5,6,8,9,12,13,15,16,21:24,28,29,30,32:34,36,37)]

cor(training, use = 'pairwise.complete.obs')
# Compute the correlation matrix and visualize it
cor.nhanes = cor(training)
cor.nhanes
corrplot(cor.nhanes, method="number")
```

#Jarque-Bera test (>3,000 Sample Size) for all numerical attributes
```{r}
JarqueBeraTest(training$hospital_stays)
JarqueBeraTest(training$time_walking_biking_perday)
JarqueBeraTest(training$moderate_activity)
JarqueBeraTest(training$frozen_meals)
JarqueBeraTest(training$bp_age)
JarqueBeraTest(training$height)
JarqueBeraTest(training$weight)
JarqueBeraTest(training$sleep_weekdays)
JarqueBeraTest(training$hours_worked)
JarqueBeraTest(training$time_outdoors_weekdays)
JarqueBeraTest(training$time_outdoors_weekends)
JarqueBeraTest(training$sunburn)
JarqueBeraTest(training$vigorous_activity)
JarqueBeraTest(training$vigorous_activity_atwork)
JarqueBeraTest(training$time_sitting)
JarqueBeraTest(training$take_away_food)
JarqueBeraTest(training$gender)
JarqueBeraTest(training$age)
JarqueBeraTest(training$ethnicity)
```

```{r}
# PCA_Plot functions - backend code

PCA_Plot = function(pcaData)
{
  library(ggplot2)
  
  theta = seq(0,2*pi,length.out = 100)
  circle = data.frame(x = cos(theta), y = sin(theta))
  p = ggplot(circle,aes(x,y)) + geom_path()
  
  loadings = data.frame(pcaData$rotation, .names = row.names(pcaData$rotation))
  p + geom_text(data=loadings, mapping=aes(x = PC1, y = PC2, label = .names, colour = .names, fontface="bold")) +
    coord_fixed(ratio=1) + labs(x = "PC1", y = "PC2")
}

PCA_Plot_Secondary = function(pcaData)
{
  library(ggplot2)
  
  theta = seq(0,2*pi,length.out = 100)
  circle = data.frame(x = cos(theta), y = sin(theta))
  p = ggplot(circle,aes(x,y)) + geom_path()
  
  loadings = data.frame(pcaData$rotation, .names = row.names(pcaData$rotation))
  p + geom_text(data=loadings, mapping=aes(x = PC3, y = PC4, label = .names, colour = .names, fontface="bold")) +
    coord_fixed(ratio=1) + labs(x = "PC3", y = "PC4")
}

PCA_Plot_Psyc = function(pcaData)
{
  library(ggplot2)
  
  theta = seq(0,2*pi,length.out = 100)
  circle = data.frame(x = cos(theta), y = sin(theta))
  p = ggplot(circle,aes(x,y)) + geom_path()
  
  loadings = as.data.frame(unclass(pcaData$loadings))
  s = rep(0, ncol(loadings))
  for (i in 1:ncol(loadings))
  {
    s[i] = 0
    for (j in 1:nrow(loadings))
      s[i] = s[i] + loadings[j, i]^2
    s[i] = sqrt(s[i])
  }
  
  for (i in 1:ncol(loadings))
    loadings[, i] = loadings[, i] / s[i]
  
  loadings$.names = row.names(loadings)
  
  p + geom_text(data=loadings, mapping=aes(x = PC1, y = PC2, label = .names, colour = .names, fontface="bold")) +
    coord_fixed(ratio=1) + labs(x = "PC1", y = "PC2")
}

PCA_Plot_Psyc_Secondary = function(pcaData)
{
  library(ggplot2)
  
  theta = seq(0,2*pi,length.out = 100)
  circle = data.frame(x = cos(theta), y = sin(theta))
  p = ggplot(circle,aes(x,y)) + geom_path()
  
  loadings = as.data.frame(unclass(pcaData$loadings))
  s = rep(0, ncol(loadings))
  for (i in 1:ncol(loadings))
  {
    s[i] = 0
    for (j in 1:nrow(loadings))
      s[i] = s[i] + loadings[j, i]^2
    s[i] = sqrt(s[i])
  }
  
  for (i in 1:ncol(loadings))
    loadings[, i] = loadings[, i] / s[i]
  
  loadings$.names = row.names(loadings)
  
  print(loadings)
  p + geom_text(data=loadings, mapping=aes(x = PC3, y = PC4, label = .names, colour = .names, fontface="bold")) +
    coord_fixed(ratio=1) + labs(x = "PC3", y = "PC4")
}

```

```{r}
#Test KMO Sampling Adequacy
library(psych)
KMO(training)

#Test Bartlett's Test of Sphericity
library(REdaS)
bart_spher(training)


#Test for Reliability Analysis using Cronbach's Alpha
library(psych)
alpha(training, check.keys = TRUE)
```


```{r}
#Create PCA, scaling the data, gives rotation matrrix
p = prcomp(training, center=T, scale=T)
#p = prcomp(responses2)
p

#Check Scree Plot
plot(p)
abline(1, 0)

#Check PCA Summary Information, info of standard deviations and breakdown of varaibles 
summary(p)
print(p)
```

#components when using standard PCA method (not final)
```{r}
p2 = psych::principal(training, rotate="varimax", nfactors=4, scores=TRUE)
p2
print(p2$loadings, cutoff=.3, sort=T)
```


Mixed PCA
###########################################################################################
```{r}

training = nhanes[, -c(1,25,26)] # remove correlated variables
training$chest_pain = factor(training$chest_pain)
training$chest_pain_walk_fast = factor(training$chest_pain_walk_fast)
training$chest_pain_normal_pace = factor(training$chest_pain_normal_pace)
training$overweight = factor(training$overweight)
training$hypertension = factor(training$hypertension)
training$gluten_free = factor(training$gluten_free)
training$stroke = factor(training$stroke)
training$diabetes_relatives = factor(training$diabetes_relatives)
training$last_doctor_visit = factor(training$last_doctor_visit)
training$diabetes = factor(training$diabetes)
training$diabetes_risk = factor(training$diabetes_risk)
training$feel_at_risk_diabetes = factor(training$feel_at_risk_diabetes)
training$smoking = factor(training$smoking)
training$gender = factor(training$gender)
training$ethnicity = factor(training$ethnicity)
training$education_level = factor(training$education_level)
training$annual_income = factor(training$annual_income)
training$bc_checked = factor(training$bc_checked)
```

#loadings and graphical representations
```{r}
split <- splitmix(training)
X1 <- split$X.quanti 
X2 <- split$X.quali 
res.pcamix <- PCAmix(X.quanti=X1, X.quali=X2,rename.level=TRUE,
                     graph=FALSE)

res.pcamix
res.pcamix$eig
?plot.PCAmix
par(mfrow=c(2,2))
plot(res.pcamix,choice="ind",coloring.ind=X2$diabetes_risk,label=FALSE,
     posleg="bottomright", main="Observations")
plot(res.pcamix,choice="levels",xlim=c(-1.5,2.5), main="Levels")
plot(res.pcamix,choice="cor",main="Numerical variables")
plot(res.pcamix,choice="sqload",coloring.var=T, leg=TRUE,
     posleg="topright", main="All variables")
```

#training/test samples
```{r}
set.seed(10)
test <- sample(1:nrow(training),2700)
train.pcamix <- PCAmix(X1[-test,],X2[-test,],ndim=25,graph=FALSE,rename.level=TRUE)
pred <- predict(train.pcamix,X1[test,],X2[test,])
head(pred)

par(mfrow=c(1,1))
plot(train.pcamix,axes=c(1,2),label=FALSE,main="Observations map")
points(pred,col=2,pch=16)
legend("bottomright",legend = c("train","test"),fill=1:2,col=1:2)
```


```{r}
data(training)
#Create the datatable
train <- training
#Split the datatable
split <- splitmix(train)
X1 <- split$X.quanti 
X2 <- split$X.quali 
res.pcamix <- PCAmix(X.quanti=X1, X.quali=X2,ndim=25, rename.level=TRUE,
                     graph=FALSE)
#In order to choose the number of dimension to rotate, we look at the inertia of the principal components
res.pcamix$eig 
res.pcarot <- PCArot(res.pcamix,dim=5,graph=FALSE)
res.pcarot$eig #variance of the rotated PCs
#before rotation
round(res.pcamix$sqload,digit=2)
res.pcarot$quali
#after rotation
round(res.pcarot$sqload,digit=2)
#remove variables with low loading values
round(res.pcarot$sqload[-c(1,4,8,9,12,15,16,18,20,23,24,26),1:5],digit=2)
```



############################################################################################
Logistic Regression Analysis
###########################################################################################
#exploratory analysis
```{r}

nhanes
dim(nhanes)
## 8,957 records, 37 variables
 
head(nhanes)
names(nhanes)
 
#check missing values
sum(is.na(nhanes))
#0 total missing values
 
# create new field for BMI
nhanes$BMI = (nhanes$height / nhanes$weight)
 
#Create new subsets of data
 
disease_risk <- nhanes[,19:32]
weight_activity_diet <- nhanes[c(1:18,38)]
demographics <- nhanes[,33:37]
 
dM<-cor(disease_risk, method="spearman")
round(dM,2)
corrplot(cor(dM,method="spearman"), method = "number", type = "lower")
 
wM<-cor(weight_activity_diet, method="spearman")
round(wM,2)
corrplot(cor(wM,method="spearman"), method = "number", type = "lower")
 
## most weighed is 0.93 highly correlated to weight
## weighed 10 years ago is 0.83 highly correlated to weight
## weight is highly correlated to BMI
## remove most weighed, weighed 10 years ago and weight
 
demoM<-cor(demographics, method="spearman")
round(demoM,2)
corrplot(cor(demoM,method="spearman"), method = "number", type = "lower")
 
nhanesCompletedv3 <- nhanes[c(1:5, 10:38)]
 
## double check correlation
totalM<-cor(nhanesCompletedv3, method="spearman")
round(totalM,2)
## correlations all seem moderate now
 
```
 
#make a copy of the enhanced data set
```{r}
#df_nhanesdataset <- nhanesCompletedv3
df_nhanesdataset <- nhanes
str(df_nhanesdataset)
sum(is.na(df_nhanesdataset))
```

#Load the necessary packages
```{r}
install.packages("dplyr")
install.packages("mosaic")
install.packages("lubridate")
install.packages("ggplot2")
 
library(dplyr)
library(mosaic)
library(lubridate)
library(ggplot2)
```

# data cleaning
```{r}
library(dplyr)
nhanescleansed <- df_nhanesdataset %>%
  transmute(diabetes_risk = as.factor(diabetes_risk)
            , sunburn = as.factor(sunburn)
            , annual_income = as.factor(annual_income)
            , bp_age = scale(bp_age)
            , bc_checked = as.factor(bc_checked)
            , smoking = as.factor(smoking)
            , hospital_stays = scale(hospital_stays)
            , last_doctor_visit = scale(last_doctor_visit)
            , time_walking_biking_perday = scale(time_walking_biking_perday)
            , moderate_activity = scale(moderate_activity)
            , chest_pain = as.factor(chest_pain)
            , chest_pain_walk_fast = as.factor(chest_pain_walk_fast)
            , chest_pain_normal_pace = as.factor(chest_pain_normal_pace)
            , overweight = as.factor(overweight)
            , hypertension = as.factor(hypertension)
            , gluten_free = as.factor(gluten_free)
            , stroke = as.factor(stroke)
            , diabetes_relatives = as.factor(diabetes_relatives)
            , diabetes = as.factor(diabetes)
            , feel_at_risk_diabetes = as.factor(feel_at_risk_diabetes)
            , ethnicity = as.factor(ethnicity)
            , education_level = as.factor(education_level)
            , moderate_activity = scale(moderate_activity)
            , frozen_meals = scale(frozen_meals)
            , age = scale(age)
            , sleep_weekdays = scale(sleep_weekdays)
            , hours_worked = scale(hours_worked)
            , time_outdoors_weekdays = scale(time_outdoors_weekdays)
            , time_outdoors_weekends = scale(time_outdoors_weekends)
            , time_sitting = scale(time_sitting)
            , take_away_food = scale(take_away_food)
            , gender = as.factor(gender)
            , vigorous_activity_atwork = scale(vigorous_activity_atwork)
            , vigorous_activity = scale(vigorous_activity)
            , BMI = scale(BMI)
  ) %>%
 
  mutate(diabetes_risk = relevel(diabetes_risk, ref = '0')
         , annual_income = relevel(annual_income, ref = '15')
         , sunburn = relevel(sunburn, ref = '2')
         , bc_checked = relevel(bc_checked, ref = '4')
         , smoking = relevel(smoking, ref = '2')
         , overweight = relevel(overweight, ref = '0')
         , chest_pain_normal_pace = relevel(chest_pain_normal_pace, ref = '0')
         , chest_pain = relevel(chest_pain, ref = '0')
         , chest_pain_walk_fast = relevel(chest_pain_walk_fast, ref = '0')
         , hypertension = relevel(hypertension, ref = '0')
         , gluten_free = relevel(gluten_free, ref = '0')
         , stroke = relevel(stroke, ref = '0')
         , diabetes_relatives = relevel(diabetes_relatives, ref = '0')
         , diabetes = relevel(diabetes, ref = '0')
         , feel_at_risk_diabetes = relevel(feel_at_risk_diabetes, ref = '0')
         , gender = relevel(gender, ref = '1')
         , ethnicity = relevel(ethnicity, ref = '3')
         , education_level = relevel(education_level, ref = '5')
  )
 
#create a matrix for annual income
incomedummies.matrx <- model.matrix(~nhanescleansed$annual_income)
 
#Convert the model matrix into a data frame
incomedummies.frame <- data.frame(incomedummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, incomedummies.frame)
 
#create a matrix for sunburn
sunburndummies.matrx <- model.matrix(~nhanescleansed$sunburn)
 
#Convert the model matrix into a data frame
sunburndummies.frame <- data.frame(sunburndummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, sunburndummies.frame)
 
#create a matrix for bc checked
bccheckeddummies.matrx <- model.matrix(~nhanescleansed$bc_checked)
 
#Convert the model matrix into a data frame
bccheckeddummies.frame <- data.frame(bccheckeddummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, bccheckeddummies.frame)
 
#create a matrix for smoking
smokingdummies.matrx <- model.matrix(~nhanescleansed$smoking)
 
#Convert the model matrix into a data frame
smokingdummies.frame <- data.frame(smokingdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, smokingdummies.frame)
 
#create a matrix for chest pain walk fast
overweightdummies.matrx <- model.matrix(~nhanescleansed$overweight)
 
#Convert the model matrix into a data frame
overweightdummies.frame <- data.frame(overweightdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, overweightdummies.frame)
 
#create a matrix for chest pain walk fast
chestpainnormaldummies.matrx <- model.matrix(~nhanescleansed$chest_pain_normal_pace)
 
#Convert the model matrix into a data frame
chestpainnormaldummies.frame <- data.frame(chestpainnormaldummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, chestpainnormaldummies.frame)
 
#create a matrix for chest pain walk fast
chestpainfastdummies.matrx <- model.matrix(~nhanescleansed$chest_pain_walk_fast)
 
#Convert the model matrix into a data frame
chestpainfastdummies.frame <- data.frame(chestpainfastdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, chestpainfastdummies.frame)
 
#create a matrix for chest pain
chestpaindummies.matrx <- model.matrix(~nhanescleansed$chest_pain)
 
#Convert the model matrix into a data frame
chestpaindummies.frame <- data.frame(chestpaindummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, chestpaindummies.frame)
 
#create a matrix for hypertension
hypertensiondummies.matrx <- model.matrix(~nhanescleansed$hypertension)
 
#Convert the model matrix into a data frame
hypertensiondummies.frame <- data.frame(hypertensiondummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, hypertensiondummies.frame)
 
#create a matrix for gluten_free
gluten_freedummies.matrx <- model.matrix(~nhanescleansed$gluten_free)
 
#Convert the model matrix into a data frame
gluten_freedummies.frame <- data.frame(gluten_freedummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, gluten_freedummies.frame)
 
#create a matrix for stroke
strokedummies.matrx <- model.matrix(~nhanescleansed$stroke)
 
#Convert the model matrix into a data frame
strokedummies.frame <- data.frame(strokedummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, strokedummies.frame)
 
#create a matrix for diabetes_relatives
diabetesrelativesdummies.matrx <- model.matrix(~nhanescleansed$diabetes_relatives)
 
#Convert the model matrix into a data frame
diabetesrelativesdummies.frame <- data.frame(diabetesrelativesdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, diabetesrelativesdummies.frame)
 
#create a matrix for diabetes
diabetesdummies.matrx <- model.matrix(~nhanescleansed$diabetes)
 
#Convert the model matrix into a data frame
diabetesdummies.frame <- data.frame(diabetesdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, diabetesdummies.frame)
 
#create a matrix for feel_at_risk_diabetes
feel_at_risk_diabetesdummies.matrx <- model.matrix(~nhanescleansed$feel_at_risk_diabetes)
 
#Convert the model matrix into a data frame
feel_at_risk_diabetesdummies.frame <- data.frame(feel_at_risk_diabetesdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, feel_at_risk_diabetesdummies.frame)
 
#create a matrix for gender
genderdummies.matrx <- model.matrix(~nhanescleansed$gender)
 
#Convert the model matrix into a data frame
genderdummies.frame <- data.frame(genderdummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, genderdummies.frame)
 
#create a matrix for ethnicity
ethnicitydummies.matrx <- model.matrix(~nhanescleansed$ethnicity)
 
#Convert the model matrix into a data frame
ethnicitydummies.frame <- data.frame(ethnicitydummies.matrx)
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, ethnicitydummies.frame)
 
#create a matrix for education_level
education_leveldummies.matrx <- model.matrix(~nhanescleansed$education_level)
 
#Convert the model matrix into a data frame
education_leveldummies.frame <- data.frame(education_leveldummies.matrx )
 
#bind the data frame to the data set
nhanescleansed <- cbind(nhanescleansed, education_leveldummies.frame)
```


 
# selecting and renaming variables
# removing height and weight
```{r}
library(dplyr)
nhanesselected <- nhanescleansed %>%
  select(diabetes_risk = diabetes_risk
          , sunburnYes = nhanescleansed.sunburn1
          , annual_income0to5 = nhanescleansed.annual_income1
         , annual_income5to10 = nhanescleansed.annual_income2
         , annual_income10to15 = nhanescleansed.annual_income3
         , annual_income15to20 = nhanescleansed.annual_income4
         , annual_income20to25 = nhanescleansed.annual_income5
         , annual_income25to35 = nhanescleansed.annual_income6
         , annual_income35to45 = nhanescleansed.annual_income7
         , annual_income45to55 = nhanescleansed.annual_income8
         , annual_income55to65 = nhanescleansed.annual_income9
         , annual_income65to75 = nhanescleansed.annual_income10
         , annual_income20plus = nhanescleansed.annual_income12
         , annual_incomeunder20 = nhanescleansed.annual_income13
         , annual_income75to100 = nhanescleansed.annual_income14
         , bccheckedlessthan1yr = nhanescleansed.bc_checked1
         , bcchecked1to2yr = nhanescleansed.bc_checked2
         , bcchecked2to5yr = nhanescleansed.bc_checked3
         , smokingeveryday = nhanescleansed.smoking1
         , smokingNo = nhanescleansed.smoking3
         , chest_painYes = nhanescleansed.chest_pain1
         , chest_pain_walk_fastYes = nhanescleansed.chest_pain_walk_fast1
         , chest_pain_walk_fastNo = nhanescleansed.chest_pain_walk_fast2
         , chest_pain_normalYes = nhanescleansed.chest_pain_normal_pace1
  , overweightYes = nhanescleansed.overweight1
  , hypertensionYes = nhanescleansed.hypertension1
  , glutenfreeYes = nhanescleansed.gluten_free1
  , strokeYes = nhanescleansed.stroke1
  , diabetesrelativesYes = nhanescleansed.diabetes_relatives1
  , diabetesNo = nhanescleansed.diabetes1
  , feelatriskdiabetesYes = nhanescleansed.feel_at_risk_diabetes1
  , MexicanAmer = nhanescleansed.ethnicity1
  , OtherHispanic = nhanescleansed.ethnicity2
  , NonHispBlack = nhanescleansed.ethnicity4
  , OtherRace = nhanescleansed.ethnicity5
  , Lessthangrade1 = nhanescleansed.education_level1
  , g9to11 = nhanescleansed.education_level2
  , hsgraduateorged = nhanescleansed.education_level3
  , somecollegeorAAdegree = nhanescleansed.education_level4
  , frozen_meals = frozen_meals
  , age = age
  , sleep_weekdays = sleep_weekdays
  , hours_worked = hours_worked
  , time_outdoors_weekdays = time_outdoors_weekdays
  , time_outdoors_weekends = time_outdoors_weekends
  , time_sitting = time_sitting
  , take_away_food = take_away_food
  , genderF = nhanescleansed.gender0
  , vigorous_activity_atwork = vigorous_activity_atwork
  , vigorous_activity = vigorous_activity
  , BMI = BMI
  , bp_age = bp_age
  , hospital_stays = hospital_stays
  , last_doctor_visit = last_doctor_visit
  , time_walking_biking_perday = time_walking_biking_perday
  , moderate_activity = moderate_activity
  )

```

 
# buid a model with all variables to see what variables from factors are
# the most significant -- filter out the other factor variables
```{r}
logistmodel <- glm(diabetes_risk ~ .
                   , family = "binomial"
                   , data=nhanesselected)
summary(logistmodel)
```
```{r}
nhanesselectedv2 <- nhanescleansed %>%
  select(diabetes_risk = diabetes_risk
         , sunburnYes = nhanescleansed.sunburn1
        
         #   , annual_income0to5 = nhanescleansed.annual_income1
         #   , annual_income5to10 = nhanescleansed.annual_income2
         #   , annual_income10to15 = nhanescleansed.annual_income3
         #   , annual_income15to20 = nhanescleansed.annual_income4
         #   , annual_income20to25 = nhanescleansed.annual_income5
         #   , annual_income25to35 = nhanescleansed.annual_income6
         #   , annual_income35to45 = nhanescleansed.annual_income7
         #   , annual_income45to55 = nhanescleansed.annual_income8
         , annual_income55to65 = nhanescleansed.annual_income9
         #   , annual_income65to75 = nhanescleansed.annual_income10
         #   , annual_income20plus = nhanescleansed.annual_income12
         #   , annual_incomeunder20 = nhanescleansed.annual_income13
         #   , annual_income75to100 = nhanescleansed.annual_income14
        
         , bccheckedlessthan1yr = nhanescleansed.bc_checked1
         #   , bcchecked1to2yr = nhanescleansed.bc_checked2
         #   , bcchecked2to5yr = nhanescleansed.bc_checked3
        
         , smokingeveryday = nhanescleansed.smoking1
         #   , smokingNo = nhanescleansed.smoking3
        
         , chest_painYes = nhanescleansed.chest_pain1
        
         #   , chest_pain_walk_fastYes = nhanescleansed.chest_pain_walk_fast1
         , chest_pain_walk_fastNo = nhanescleansed.chest_pain_walk_fast2
        
         , chest_pain_normalYes = nhanescleansed.chest_pain_normal_pace1
        
         , overweightYes = nhanescleansed.overweight1
        
         , hypertensionYes = nhanescleansed.hypertension1
        
         , glutenfreeYes = nhanescleansed.gluten_free1
        
         , strokeYes = nhanescleansed.stroke1
        
         , diabetesrelativesYes = nhanescleansed.diabetes_relatives1
        
         , diabetesNo = nhanescleansed.diabetes1
        
         , feelatriskdiabetesYes = nhanescleansed.feel_at_risk_diabetes1
        
         #   , MexicanAmer = nhanescleansed.ethnicity1
         #   , OtherHispanic = nhanescleansed.ethnicity2
         , NonHispBlack = nhanescleansed.ethnicity4
         #   , OtherRace = nhanescleansed.ethnicity5
        
         #   , Lessthangrade1 = nhanescleansed.education_level1
         #   , g9to11 = nhanescleansed.education_level2
         , hsgraduateorged = nhanescleansed.education_level3
         #   , somecollegeorAAdegree = nhanescleansed.education_level4
        
         , frozen_meals = frozen_meals
         , age = age
         , sleep_weekdays = sleep_weekdays
         , hours_worked = hours_worked
         , time_outdoors_weekdays = time_outdoors_weekdays
         , time_outdoors_weekends = time_outdoors_weekends
         , time_sitting = time_sitting
         , take_away_food = take_away_food
        
         , genderF = nhanescleansed.gender0
        
         , vigorous_activity_atwork = vigorous_activity_atwork
         , vigorous_activity = vigorous_activity
         , BMI = BMI
        
         , bp_age = bp_age
         , hospital_stays = hospital_stays
         , last_doctor_visit = last_doctor_visit
         , time_walking_biking_perday = time_walking_biking_perday
         , moderate_activity = moderate_activity
         ,
  )
 
```
 
#build a model with new dataset
```{r}
logistmodel2 <- glm(diabetes_risk ~ .
                   , family = "binomial"
                   , data=nhanesselectedv2)
summary(logistmodel2)
```
 
# remove insignificant variables and re-run model
```{r}
nhanesselectedv3 <- nhanescleansed %>%
  select(diabetes_risk = diabetes_risk
     ##    , sunburnYes = nhanescleansed.sunburn1
        
         #   , annual_income0to5 = nhanescleansed.annual_income1
         #   , annual_income5to10 = nhanescleansed.annual_income2
         #   , annual_income10to15 = nhanescleansed.annual_income3
         #   , annual_income15to20 = nhanescleansed.annual_income4
         #   , annual_income20to25 = nhanescleansed.annual_income5
         #   , annual_income25to35 = nhanescleansed.annual_income6
         #   , annual_income35to45 = nhanescleansed.annual_income7
         #   , annual_income45to55 = nhanescleansed.annual_income8
         , annual_income55to65 = nhanescleansed.annual_income9
         #   , annual_income65to75 = nhanescleansed.annual_income10
         #   , annual_income20plus = nhanescleansed.annual_income12
         #   , annual_incomeunder20 = nhanescleansed.annual_income13
         #   , annual_income75to100 = nhanescleansed.annual_income14
        
      #   , bccheckedlessthan1yr = nhanescleansed.bc_checked1
         #   , bcchecked1to2yr = nhanescleansed.bc_checked2
         #   , bcchecked2to5yr = nhanescleansed.bc_checked3
        
         , smokingeveryday = nhanescleansed.smoking1
         #   , smokingNo = nhanescleansed.smoking3
        
      #   , chest_painYes = nhanescleansed.chest_pain1
        
         #   , chest_pain_walk_fastYes = nhanescleansed.chest_pain_walk_fast1
         , chest_pain_walk_fastNo = nhanescleansed.chest_pain_walk_fast2
        
         , chest_pain_normalYes = nhanescleansed.chest_pain_normal_pace1
        
         , overweightYes = nhanescleansed.overweight1
        
     #    , hypertensionYes = nhanescleansed.hypertension1
        
         , glutenfreeYes = nhanescleansed.gluten_free1
        
         , strokeYes = nhanescleansed.stroke1
        
         , diabetesrelativesYes = nhanescleansed.diabetes_relatives1
        
         , diabetesNo = nhanescleansed.diabetes1
        
         , feelatriskdiabetesYes = nhanescleansed.feel_at_risk_diabetes1
        
         #   , MexicanAmer = nhanescleansed.ethnicity1
         #   , OtherHispanic = nhanescleansed.ethnicity2
         , NonHispBlack = nhanescleansed.ethnicity4
         #   , OtherRace = nhanescleansed.ethnicity5
        
         #   , Lessthangrade1 = nhanescleansed.education_level1
         #   , g9to11 = nhanescleansed.education_level2
         , hsgraduateorged = nhanescleansed.education_level3
         #   , somecollegeorAAdegree = nhanescleansed.education_level4
        
         , frozen_meals = frozen_meals
         , age = age
    #     , sleep_weekdays = sleep_weekdays
    #     , hours_worked = hours_worked
   #      , time_outdoors_weekdays = time_outdoors_weekdays
    #     , time_outdoors_weekends = time_outdoors_weekends
     #    , time_sitting = time_sitting
   #      , take_away_food = take_away_food
        
         , genderF = nhanescleansed.gender0
        
         , vigorous_activity_atwork = vigorous_activity_atwork
      #   , vigorous_activity = vigorous_activity
         , BMI = BMI
        
         , bp_age = bp_age
    #     , hospital_stays = hospital_stays
         , last_doctor_visit = last_doctor_visit
    #     , time_walking_biking_perday = time_walking_biking_perday
    #     , moderate_activity = moderate_activity
         ,
  )
 
```
# build new model
#### final model for analysis
```{r}
logistmodel3 <- glm(diabetes_risk ~ .
                    , family = "binomial"
                    , data=nhanesselectedv3)
summary(logistmodel3)
 
```
# odds ratios only
```{r}
exp(coef(logistmodel3))
```
 
#run VIF
```{r}
VIF(logistmodel3)
# no multicollinearity
```
 
#check automatic models
```{r}
null = glm(diabetes_risk ~ 1, data=nhanesselectedv3, family = "binomial")
null
 
full = glm(diabetes_risk ~ ., data=nhanesselectedv3, family = "binomial")
full
```
#Forward Regression
```{r}
forwardnhanescleased = step(null, scope = list(lower=null, upper=full), direction="forward")
summary(forwardnhanescleased)
 
summary(logistmodel3)
```
 
 
#Backward Regression
```{r}
backwardnhanescleased = step(full, direction="backward")
summary(backwardnhanescleased)
```
 
#Stepwise Regression
```{r}
stepnhanescleased = step(null, scope = list(upper=full), direction="both")
summary(stepnhanescleased)
```

#######################################################################################################################################################################
#CCA
######################################################################################################################################################################

#Assuming that we linearity in our model
#Assuming that there is no multicollinearity
#Assuming that we have no normality 
# We should have relatively equal variance
#####################################################################################
####################################################################################
#Canonical Correlation

#diabetes variables of interest
```{r}
library(dplyr)
diabetesRisk <- nhanes %>%
  select(diabetes_relatives = diabetes_relatives
         , diabetes_risk = diabetes_risk
         , feel_at_risk_diabetes = feel_at_risk_diabetes
         #, stroke = stroke
         )
```

#select activityLevel variables of interest
```{r}
activitylevel <- nhanescleansed %>%
  select(time_walking_biking_perday = time_walking_biking_perday
         , moderate_activity = moderate_activity
         , time_outdoors_weekdays = time_outdoors_weekdays
         , time_outdoors_weekends = time_outdoors_weekends
         , vigorous_activity = vigorous_activity
         )%>%
  mutate(time_walking_biking_perday = scale(time_walking_biking_perday)
         , moderate_activity = scale(moderate_activity)
         , time_outdoors_weekdays = scale(time_outdoors_weekdays)
         , time_outdoors_weekends = scale(time_outdoors_weekends)
         , vigorous_activity = scale(vigorous_activity))
```



#select variables for demographic
```{r}
demographic <- nhanes %>%
  select(height = height
          , weight = weight
         , gender = gender
         , education_level = education_level
         , ethnicity = ethnicity
         , age = age, annual_income = annual_income, weight_10yearsago = weight_10yearsago#, hours_worked = hours_worked
         
         ) %>%
  mutate(BMI = scale(weight)/scale(height)
         , age = scale(age)
         , annual_income = scale(annual_income)
         , weight_10yearsago = scale(weight_10yearsago)
         )
demographic$height <- NULL
demographic$weight <- NULL
```

#select the variables for healthcondition
```{r}
healthconditions <- nhanes %>%
  select(hypertension = hypertension
         , bc_checked = bc_checked
         , age = age
         , smoking = smoking
         , overweight = overweight
         , stroke = stroke
         , chest_pain = chest_pain
         #, BMI = weight/height
         )
```


###################################################################
# Exploring correlations between psychology and academics


###################################################################
# This is a nice function for computing the Wilks lambdas for 
# CCA data from the CCA library's method
# It computes the wilkes lambas the degrees of freedom and te 
# p-values
###################################################################
```{r}
ccaWilks = function(set1, set2, cca)
{
  ev = ((1 - cca$cor^2))
  ev
  
  n = dim(set1)[1]
  p = length(set1)
  q = length(set2)
  k = min(p, q)
  m = n - 3/2 - (p + q)/2
  m
  
  w = rev(cumprod(rev(ev)))
  
  # initialize
  d1 = d2 = f = vector("numeric", k)
  
  for (i in 1:k) 
  {
    s = sqrt((p^2 * q^2 - 4)/(p^2 + q^2 - 5))
    si = 1/s
    d1[i] = p * q
    d2[i] = m * s - p * q/2 + 1
    r = (1 - w[i]^si)/w[i]^si
    f[i] = r * d2[i]/d1[i]
    p = p - 1
    q = q - 1
  }
  
  pv = pf(f, d1, d2, lower.tail = FALSE)
  dmat = cbind(WilksL = w, F = f, df1 = d1, df2 = d2, p = pv)
}
```

################################################################
# using the yacca package
################################################################
```{r}
options(scipen = 0)
library(yacca)
c2 = cca(demographic, diabetesRisk)

summary(c2)
```
#Helio plot for CV1
```{r}
helio.plot(c2
           , cv=1
           , x.name="Demographic"
           , y.name = "Diabetes Risk"
           #, sub = ""
           )
```

#Helio plot for CV2
```{r}
helio.plot(c2
           , cv=2
           , x.name="Diabetes values"
           , y.name = "Demographic values"
           )
```

##################################################################################################
#Correspondence analysis
##################################################################################################
#select the required varaibles of interest for Correspondence analysis
```{r}
nhanescleansedCat <- nhanescleansed %>%
  select(gender = gender
         , education_level = education_level
         , ethnicity = ethnicity
         , stroke = stroke)
```

#structure
```{r}
str(nhanescleansedCat)
```






