---
title: "Micedatasetexample"
author: "Ronaldlee Ejalu"
date: "2/17/2021"
output: html_document
---

#load the necessary packages
```{r}
library(RCurl, quietly = TRUE)
library(mice, quietly = TRUE)
```

# set working directory
```{r}
setwd("C:/Users/rejalu1/OneDrive - Henry Ford Health System/DSC424/HomeWork3")
```


#make a copy of the nhanes dataset
```{r}
df_nhanes <- nhanes
```


#Check data for missing values
```{r}
sapply(df_nhanes
       , function(x) sum(is.na(x))
       )
```

```{r}
str(nhanes)
```

#transform variables into numeric or or factors 
#depending on the data type
```{r}
library(dplyr)
df_nhanes <- df_nhanes %>%
  mutate(age = as.factor(age)
         , bmi = as.numeric(bmi)
         , hyp = as.numeric(hyp)
         , chl = as.numeric(chl)
  )
```

#it is now time to impute the data.
```{r}
library(mice, quietly = TRUE)
init=mice(df_nhanes, maxit = 0)
meth=init$method
predM=init$predictorMatrix
```
#To impute the missing values, mice package use an algorithm in a #such a way that use information from other variables in the dataset #to predict and impute the missing values. Therefore, you may not #want to use a certain variable as predictors. For example, the ID #variable does not have any predictive value.

#The code below will remove the variable as a predictor but still #will be imputed. Just for illustration purposes, I select the BMI #variable to not be included as predictor during imputation
```{r}
#predM[, c("age")] = 0
predM[, c("bmi")] = 0
```

#Now let specify the methods for imputing the missing values. There #are specific methods for continous, binary and ordinal variables. I #set different methods for each variable. You can add more than one #variable in each method.
```{r}
meth[c("hyp")] = "norm"
meth[c("chl")] = "norm"
meth[c("bmi")] = "norm"
```

#Now it is time to run the multiple (m=5) imputation.
```{r}
set.seed(103)
imputed=mice(df_nhanes
             , method=meth
             , predictorMatrix=predM
             , m=5
)
```
#Create a dataset after imputation.
```{r}
imputed <- mice::complete(imputed)
```
#Check for missing values in the imputed dataset
```{r}
sapply(imputed
       , function(x) sum(is.na(x))
       )
```

#############Checking for accuracy#############
```{r}
#chl
#actual <- nhanes$chl[is.na(df_nhanes$chl)]
#predicted <- imputed$chl[is.na(df_nhanes$chl)]

#hyp
actual <- nhanes$hyp[is.na(df_nhanes$hyp)]
predicted <- imputed$hyp[is.na(df_nhanes$hyp)]

#table(actual)
#table(predicted)

#mean(actual)
mean(predicted)
```
